---
title: 'Stop and Frisk in DC: Jul14-Aug18 2019'
author: "Brendan Freehart"
date: "October 14, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 1
    toc_float: true
    theme: spacelab
    highlight: tango
    fig_caption: yes
---
_This site was built using RMarkdown_

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(RColorBrewer)
library(knitr)
library(kableExtra)
library(plotly)
library(mapview)
library(tidyverse)

load('frisk_19_clean.RData')

frisks_per_hour <- nrow(frisks) / (as.integer(max(frisks$stop_date) - min(frisks$stop_date))*24)
semicolon_delim_pattern <- "(.+)\\;\\s+(.+)\\;\\s+(.+)\\;\\s+(.+)\\;\\s+(.+)|(.+)\\;\\s+(.+)\\;\\s+(.+)\\;\\s+(.+)|(.+)\\;\\s+(.+)\\;\\s+(.+)|(.+)\\;\\s+(.+)|(.+)"

```

# Stop and Frisks in DC
Stop and Frisk is the practice of temporarily detaining, questioning, and searching civilians on the street without probable cause [^1]. Washington, DC's Metropolitan Police Department is required by the NEAR Act to record Stop and Frisks [^2].

Between July 14th and August 18th, there were `r sprintf("%.1f", frisks_per_hour)` stop and frisks per hour.

```{r top level}
frisks %>% 
  summarise(
    stops = n(), 
    arrests = sum(!is.na(arrest_date)),
    prcnt_arrests = arrests / stops
  ) %>%
  mutate(
    prcnt_arrests = sprintf("%.1f%%", prcnt_arrests*100)
  ) %>% 
  kable(col.names = c("Stop and Frisks","Arrests","% of Frisks leading to Arrest")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", position = "right", full_width = F))

```


***
# Who is Targeted by Stop and Frisk?

Reported Stop and Frisk incidents show a substantial bias towards young, black men. **93% of people stopped were black**.

```{r by race}
gg_race <- frisks %>%
  group_by(race_ethnicity) %>% 
  summarise(
    Arrest = sum(!is.na(arrest_date)),
    `No Arrest` = n() - Arrest
  ) %>% 
  gather(key = 'Outcome', value = 'Stops', -race_ethnicity) %>% 
  ggplot(aes(reorder(race_ethnicity, -Stops), Stops, fill = Outcome)) +
  geom_bar(stat = "identity") + 
  theme_minimal(base_family="Montserrat") + 
  theme(axis.text.x = element_text(angle=60, hjust=1), axis.title.x = element_blank(), plot.title = element_text(size = 24)) + 
  scale_fill_manual(values = rev(brewer.pal(4,"Blues"))) + 
  labs(title = '# of Frisks by Recorded Race / Ethnicity', y = 'Stop and Frisks')

gg_race %>% 
  ggplotly(tooltip = "Stops") %>% 
  add_annotations(
    yref="paper", 
    xref="paper", 
    y=1.15, 
    x=0, 
    text="# of Frisks by Recorded Race / Ethnicity", 
    showarrow=F, 
    font=list(size=24),
    hovertemplate = '%{y} Stops'
  ) %>% 
  layout(
    title=FALSE
  )

```


```{r demog table}
gg_demo <- frisks %>% 
  mutate(
    demo_segment = paste(race_ethnicity,gender,age_group)
  ) %>% 
  group_by(demo_segment) %>% 
  summarise(
    prcnt_of_total_stops = n() / nrow(frisks),
    arrests = sum(!is.na(arrest_date)),
    non_arrests = n() - arrests
  ) %>% 
  ungroup() %>% 
  mutate(
    demo_segment = if_else(prcnt_of_total_stops < 0.01, 'Other', demo_segment)
  ) %>% 
  group_by(demo_segment) %>%  
  summarise(
    Arrests = sum(arrests),
    `No Arrests` = sum(non_arrests)
  ) %>% 
  gather(key = 'Outcome', value = 'Stops', -demo_segment) %>% 
  ggplot(aes(reorder(demo_segment, -Stops), Stops, fill = Outcome)) + 
  geom_bar(stat = "identity") + 
  theme_minimal(base_family="Montserrat") + 
  theme(axis.text.x = element_text(angle=60, hjust=1), axis.title.x = element_blank(), plot.title = element_text(size = 24)) + 
  scale_fill_manual(values = rev(brewer.pal(4,"Blues"))) + 
  labs(title = '# of Frisks by Race, Gender, and Age', y = 'Stop and Frisks')

gg_demo %>% 
  ggplotly(tooltip = "Stops") %>% 
  add_annotations(
    yref="paper", 
    xref="paper", 
    y=1.15, 
    x=0, 
    text="# of Frisks by Race, Gender, and Age", 
    showarrow=F, 
    font=list(size=24),
    hovertemplate = '%{y} Stops'
  ) %>% 
  layout(
    title=FALSE
  )

```

***

# What Reason is Provided for Stop and Frisks?

```{r reasons cited}
gg_reasons <- frisks %>% 
  transmute(
    reason_col1 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,2],
    reason_col2 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,3],
    reason_col3 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,4],
    reason_col4 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,5],
    reason_col5 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,6],
    reason_col6 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,7],
    reason_col7 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,8],
    reason_col8 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,9],
    reason_col9 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,10],
    reason_col10 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,11],
    reason_col11 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,12],
    reason_col12 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,13],
    reason_col13 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,14],
    reason_col14 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,15],
    reason_col15 = str_match(stop_reason_nonticket, semicolon_delim_pattern)[,16]
  ) %>% 
  gather() %>% 
  filter(!is.na(value)) %>%
  mutate(value = if_else(value=='NULL', 'Not provided', value)) %>% 
  select(reason = value) %>% 
  group_by(reason) %>% 
  summarise(cited = n()) %>%
  ggplot(aes(reorder(reason, -cited), cited), fill=reason) +
  geom_bar(stat = "identity") + 
  theme_minimal(base_family="Montserrat") + 
  theme(axis.text.x = element_text(angle=60, hjust=1), axis.title.x = element_blank(), plot.title = element_text(size = 24)) + 
  scale_fill_manual(values = rev(brewer.pal(4,"Blues"))) + 
  labs(title = 'Cited Reason for Stop and Frisk', y = 'Cited')

gg_reasons %>% 
  ggplotly(tooltip = "Cited") %>% 
  add_annotations(
    yref="paper", 
    xref="paper", 
    y=1.15, 
    x=0, 
    text="Cited Reason for Stop and Frisk", 
    showarrow=F, 
    font=list(size=24),
    hovertemplate = '%{y} Cited'
  ) %>% 
  layout(
    title=FALSE
  )

```

\

# What Charges Are Brought?
```{r arrests}
frisks %>% 
  filter(!is.na(arrest_date)) %>% 
  group_by(arrest_charges) %>% 
  summarise(
    stops = n()
  ) %>% 
  filter(stops >= 10 ) %>% 
  arrange(-stops) %>% 
  kable(col.names = c("Charges","# of Stops")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", position = "right", full_width = F))

```

*** 

# Where is Stop and Frisk happening?

```{r population}
census_with_frisks %>% 
  mapview(zcol = "Total", legend = TRUE)

```

\

```{r income}
census_with_frisks %>% 
  mapview(zcol = "MedianIncome", legend = TRUE)

```

\

```{r percentage black}
census_with_frisks %>% 
  mapview(zcol = "prcnt_black", legend = TRUE)

```

\

```{r stops}
census_with_frisks %>% 
  mapview(zcol = "total_stops", legend = TRUE)

```

\

```{r percentage of black targets}
census_with_frisks %>% 
  mapview(zcol = "prcnt_black_stops", legend = TRUE)

```

\

```{r calls for service}
census_with_frisks %>% 
  mapview(zcol = "prcnt_call_for_service", legend = TRUE)

```


***

# See also
This excellent [analysis of _reported_ DC Stop and Frisk from 2010-2017](https://gwarrenn.github.io/Stop-and-Frisk-DC/).

[^1]: Source: [Stop and Frisk Report 2013](https://web.archive.org/web/20170211080746/http://oag.state.ny.us/sites/default/files/pdfs/bureaus/civil_rights/stp_frsk.pdf)
[^2]: Source: [Implementing the NEAR Act to reduce violence in D.C.](https://www.dcpolicycenter.org/publications/implementing-near-act-reduce-violence-d-c/)


